---
title: "get_data"
author: "Xiaohao Yang"
vignette: >
  %\VignetteIndexEntry{greenSD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Donwload multi-band data from Greenspace Seasonality Data Cube with specified area/point of inerest
This may take more than 5 minutes.

### By bounding box

```{r eval=FALSE}
gs <- greenSD::get_gsds_data(bbox = c(-83.272828,42.343950,-83.218926,42.379719), year = 2022, mask = TRUE)
```

### By place name

```{r eval=FALSE}
gs <- greenSD::get_gsds_data(place = 'Detroit', year = 2022)
```

### By coordinates (point)

```{r eval=FALSE}
gs <- greenSD::get_gsds_data(location = c(-83.10215 42.38342), year = 2022)
```

### By UID and time range

```{r eval=FALSE}
# greenSD::check_available_cities()
gs <- greenSD::get_gsds_data(UID = 1825, year = 2022, time = c("03-01", "09-01"))
```

# Extract values from Greenspace Seasonality Data Cube with samples
You can extract seasonal greenspace values at multiple point locations 
within a city boundary.

```{r eval=FALSE}
boundary <- greenSD::check_urban_boundary(uid = 1825, plot = FALSE)
samples <- sf::st_sample(boundary, size = 50)
gs_samples <- greenSD::sample_values(samples, year = 2022)
```

# Visualize Seasonal Greenspace Dynamics as an Animated GIF
The `to_gif()` function converts a multi-band raster (e.g., greenspace bands 
across the growing season) into an animated GIF for quick visual exploration.

```{r eval=FALSE}
# Load example data (or use `gs` from previous step)
sample_data <- terra::rast(system.file("extdata", "detroit_gs.tif", package = "greenSD"))

# Generate GIF
gif <- greenSD::to_gif(
  r = sample_data,
  fps = 5,
  width = 600,
  height = 600,
  axes = FALSE,
  title_prefix = paste("greenspace - Day", 1:terra::nlyr(sample_data) * 10)
)

# Display in RStudio Viewer or save
print(gif)

# To save the GIF manually:
magick::image_write(gif, "greenspace_animation.gif")
```



